---
title: "Pooled Testing Strategies with Quantitative Assays "
author: "Tao Liu, PhD, Brown University"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{QuantPooledTesting}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

**Type** Package

**Title** Quantitative Pooled Testing

**Version** 1.0

**Date** "`r Sys.Date()`"

**Authors** Wanning Su, Tao Liu

**Maintainer** Wanning (wanning_su@brown.edu)

**Description** This package implements pooled testing with quantitative assays. 

**License** MIT

**URL**  https://github.com/whitneysu/QuantitativePooledTesting

**Depends** R (>= 3.2.0)

**LazyData** true

<!-- START doctoc -->
<!-- END doctoc -->


###ecdf_pool

*empirical cdf calculation*

####Description

> Generates an empirical cdf for quantitative pooled testing 

####Usage

> ecdf_pool(v, N = length(v), cutoff = max(v), ...)

####Arguments

> - **v** &nbsp;&nbsp;&nbsp;&nbsp; the numeric vector of the observations for calculating ecdf.For quantitative pooling strategies, only CDF on the support of v<cutoff is needed. 
> - **N**  &nbsp;&nbsp;&nbsp;&nbsp;  the number of observations.
> - **cutoff** cutoff value of the quantitative assay that defines test positivity. 

####Value

> a list of two items: (1) a matrix and (2) a numeric value. The matrix has 3 columns: the support consisting of values of such that v < cutoff, the empirical PMF, and the empirical CDF. The second item in the list output is a numeric value that represents the proportion of total observations that are less than cutoff value v. 
 
####Author(s)

> Wanning Su (wanning_su@brown.edu), Tao Liu, Joseph Hogan

####References

> Liu T, Hogan JW, Daniels MJ, et al. (2017). "Improved HIV-1 viral load monitoring capacity using pooled testing with marker-assisted deconvolution." *J Acquir Immune Defic Syndr* 75(5):580-587.


####Examples

```{r,eval=FALSE}
ecdf_pool(round(runif(100, 0, 20)), cutoff = 18)
```

_____


###to_ecdf_class

*Create a ecdf class for ecdf_pool function output*

####Description

> Create an ecdf class object. 

####Usage

> to_ecdf_class(x)

####Arguments

> - **x** &nbsp;&nbsp;&nbsp;&nbsp; Output object from the \code{ecdf_pool()} function 

####Value

> An object of class ecdf
 
####Author(s)

> Wanning Su (wanning_su@brown.edu), Tao Liu, Joseph Hogan

####References

> Liu T, Hogan JW, Daniels MJ, et al. (2017). "Improved HIV-1 viral load monitoring capacity using pooled testing with marker-assisted deconvolution." *J Acquir Immune Defic Syndr* 75(5):580-587.


####Examples

```{r,eval=FALSE}
to_ecdf_class(ecdf_pool(round(runif(100, 0, 20)), cutoff = 18))
```

_____

###convol.dens

*convoluting two empirical distributions*

####Description

> This function calculates the convolution of two empirical distributions up to the value of \code{cutoff} (on the sum of two random variables).

####Usage

> convol.dens(ecdf1, ecdf2, N1, N2, cutoff)


####Arguments

> - **ecdf1** &nbsp;&nbsp;&nbsp;&nbsp; An ecdf that is the matrix output object from the function \code{convol.dens()} itself or from the function
\code{ecdf_pool()}.
> - **ecdf2**  &nbsp;&nbsp;&nbsp;&nbsp;  An ecdf that is the matrix output object from the function \code{convol.dens()} itself or from the function
\code{ecdf_pool()}.
> - **N1** &nbsp;&nbsp;&nbsp;&nbsp; The size of the empirical support of \code{ecdf1}.
> - **N2**  &nbsp;&nbsp;&nbsp;&nbsp; The size of the empirical support of \code{ecdf2}.
> - **cutoff**  &nbsp;&nbsp;&nbsp;&nbsp; Cutoff of the support of the resulting convolution distribution.


####Value

> Returns a list of two items: (1) a matrix and (2) a numeric value. The matrix has 3 columns: the support consisting of values of such that v < cutoff, the empirical PMF, and the empirical CDF. The second item in the list output is a numeric value that represents the proportion of total observations that are less than cutoff value v. 
 
 
####Author(s)

> Wanning Su (wanning_su@brown.edu), Tao Liu, Joseph Hogan

####References

> Liu T, Hogan JW, Daniels MJ, et al. (2017). "Improved HIV-1 viral load monitoring capacity using pooled testing with marker-assisted deconvolution." *J Acquir Immune Defic Syndr* 75(5):580-587.


####Examples

```{r,eval=FALSE}
ecdf1 <- ecdf_pool(round(runif(100, 0, 20)), cutoff = 18)$out
ecdf2 <- ecdf_pool(round(runif(100, 0, 20)), cutoff = 17)$out
cutoff <- 18 
convol.dens(ecdf1, ecdf2, N1, N2, cutoff)
```
_____

###foo_atr

*Intermediate Function used by MPA*


####Description

> This function allows you to compute the calculate the ATR for MPA for a range of pool sizes (from 2 to pool size specified)

####Usage

> foo_atr(v, N = length(v), max_K = 5, cutoff = 1000, quietly = TRUE)


####Arguments

> - **v** &nbsp;&nbsp;&nbsp;&nbsp; the numeric vector of the observations for calculating ecdf.For quantitative pooling strategies, only CDF on the support of v<cutoff is needed. 
> - **N**  &nbsp;&nbsp;&nbsp;&nbsp;  the number of observations.
> - **cutoff** cutoff value of the quantitative assay that defines test positivity. 
> - **max_K** &nbsp;&nbsp;&nbsp;&nbsp; Maximum pool size that is under consideration.
> - **quietly**  &nbsp;&nbsp;&nbsp;&nbsp; Logical value; whether the results print to screen or not.

 
####Value

> Returns a three column matrix of 1)pool size, 2)convolution CDF, and the 3)ATR for MPA.

####Author(s)

> Wanning Su (wanning_su@brown.edu), Tao Liu, Joseph Hogan

####References

> Liu T, Hogan JW, Daniels MJ, et al. (2017). "Improved HIV-1 viral load monitoring capacity using pooled testing with marker-assisted deconvolution." *J Acquir Immune Defic Syndr* 75(5):580-587.

####Examples

```{r,eval=FALSE}
v = c(200, 1500, 1900, 800, 950)
max_K = 9
cutoff = 2000
foo_atr(v, N = length(v), max_K, cutoff, quietly = TRUE)
```

_____

###mp_atr

*ATR required by MiniPooling(MP)*


###Description

> This function calculates the ATR needed by MP for a range of pool sizes (from 2 to a pool size specified).

####Usage

> mp_atr(v, N = length(v), max_K = 5, cutoff = 1000, quietly = TRUE)

####Arguments

> - **v** &nbsp;&nbsp;&nbsp;&nbsp; the numeric vector of the observations for calculating ecdf.For quantitative pooling strategies, only CDF on the support of v<cutoff is needed. 
> - **N**  &nbsp;&nbsp;&nbsp;&nbsp;  the number of observations.
> - **cutoff** cutoff value of the quantitative assay that defines test positivity. 
> - **max_K** &nbsp;&nbsp;&nbsp;&nbsp; Maximum pool size that is under consideration.
> - **quietly**  &nbsp;&nbsp;&nbsp;&nbsp; Logical value; whether the results print to screen or not.

####Value

> The function returns a matrix of two columns: The pool size from 2:max_K and the Average Tests Required by MP given that pool size (MP_ATR).

####Author(s)

> Wanning Su (wanning_su@brown.edu), Tao Liu, Joseph Hogan

####References

> Liu T, Hogan JW, Daniels MJ, et al. (2017). "Improved HIV-1 viral load monitoring capacity using pooled testing with marker-assisted deconvolution." *J Acquir Immune Defic Syndr* 75(5):580-587.


####Examples

```{r,eval=FALSE}
v = c(200, 1500, 1900, 800, 950)
max_K = 9
cutoff = 2000
mp_atr(v, N = length(v), max_K, cutoff, quietly = TRUE)
```

_____

###mpa_atr

*ATR required by MPA*

###Description

> This function calculates the ATR needed by MP for a range of pool sizes (from 2 to a pool size specified).

####Usage

> mpa_atr(v, N = length(v), max_K = 5, cutoff = 1000, quietly = TRUE)

####Arguments

> - **v** &nbsp;&nbsp;&nbsp;&nbsp; the numeric vector of the observations for calculating ecdf.For quantitative pooling strategies, only CDF on the support of v<cutoff is needed. 
> - **N**  &nbsp;&nbsp;&nbsp;&nbsp;  the number of observations.
> - **cutoff** cutoff value of the quantitative assay that defines test positivity. 
> - **max_K** &nbsp;&nbsp;&nbsp;&nbsp; Maximum pool size that is under consideration.
> - **quietly**  &nbsp;&nbsp;&nbsp;&nbsp; Logical value; whether the results print to screen or not.

####Value

> The function returns a matrix of two columns: The pool size from 2:max_K and the Average Tests Required by MPA given that pool size (MPA_ATR).

####Author(s)

> Wanning Su (wanning_su@brown.edu), Tao Liu, Joseph Hogan

####References

> Liu T, Hogan JW, Daniels MJ, et al. (2017). "Improved HIV-1 viral load monitoring capacity using pooled testing with marker-assisted deconvolution." *J Acquir Immune Defic Syndr* 75(5):580-587.


####Examples

```{r,eval=FALSE}
v = c(200, 1500, 1900, 800, 950)
max_K = 9
cutoff = 2000
mpa_atr(v, N = length(v), max_K, cutoff, quietly = TRUE)
```

_____

###mmpa_atr

*ATR required by MMPA*

###Description

> This function calculates the ATR needed by MMPA for a range of pool sizes (from 2 to a pool size specified).

####Usage

> mmpa_atr(v, N = length(v), max_K = 5, cutoff = 1000, quietly = TRUE, s = )

####Arguments

> - **v** &nbsp;&nbsp;&nbsp;&nbsp; the numeric vector of the observations for calculating ecdf.For quantitative pooling strategies, only CDF on the support of v<cutoff is needed. 
> - **N**  &nbsp;&nbsp;&nbsp;&nbsp;  the number of observations.
> - **cutoff** cutoff value of the quantitative assay that defines test positivity. 
> - **max_K** &nbsp;&nbsp;&nbsp;&nbsp; Maximum pool size that is under consideration.
> - **quietly**  &nbsp;&nbsp;&nbsp;&nbsp; Logical value; whether the results print to screen or not.
> - **s**  &nbsp;&nbsp;&nbsp;&nbsp; Risk score that is used for ranked testing


####Value

> The function returns a matrix of two columns: The pool size from 2:max_K and the Average Tests Required by MMPA given that pool size (MMPA_ATR).

####Author(s)

> Wanning Su (wanning_su@brown.edu), Tao Liu, Joseph Hogan

####References

> Liu T, Hogan JW, Daniels MJ, et al. (2017). "Improved HIV-1 viral load monitoring capacity using pooled testing with marker-assisted deconvolution." *J Acquir Immune Defic Syndr* 75(5):580-587.


####Examples

```{r,eval=FALSE}
v = c(200, 1500, 1900, 800, 950)
max_K = 9
cutoff = 2000
s = c(5, 3, 1, 2, 4)
mmpa_atr(v, s, max_K, cutoff = 2000, quietly = TRUE)
```

